Deep ordinal classification based on cumulative link models

### 2019-10-08 paper/elsarticle-template.tex
Paper neuro

\section{Introduction}
\label{sect:introduction}
Deep learning, introduced by Lecun~et al. \cite{lecun2015deep}, combines multiple machine learning techniques and allows computational models that are composed of numerous processing layers to learn representations of data with various levels of abstraction. These methods have dramatically improved the state-of-the-art in many domains, such as image classification~\cite{cirecsan2012multi,he2016deep,krizhevsky2012imagenet}, speech recognition~\cite{hinton2012deep}, control problems~\cite{mnih2015human}, object detection~\cite{jiang2016speed,girshick2014rich}, privacy and security protection~\cite{yu2017iprivacy,yuan2019adversarial,tan2018deep}, recovery of human pose~\cite{hong2015multimodal}, semantic segmentation~\cite{long2015fully} and image retrieval~\cite{li2015weakly,jin2018deep}. Convolutional Neural Networks (CNN) are one of the types of deep networks that are designed to process data that come in the form of multiple arrays. CNNs are appropriate for images, video, speech and audio processing, and they have been used extensively in the last years for automatic classification tasks~\cite{dong2014learning,sun2013deep,ronneberger2015u}. For images, each colour channel is represented by a 2D array, and convolutional layers extract the main features from the pixels, and, after that, a fully connected layer classify every sample based on its extracted features. At the output of the CNN, a softmax function provides the probabilities of the set of classes predefined in the model for classification tasks. However, the softmax may not be the best option depending on the problem considered.

@@

\begin{figure}[!t]
	\centering
	\begin{tabular}{cccc}
		\includegraphics[width=2.4cm]{img/retinopathy/1.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/2.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/3.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/4.jpg}\\
		\includegraphics[width=2.4cm]{img/retinopathy/5.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/6.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/7.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/8.jpg}
	\end{tabular}						
	\caption{Examples taken from the Diabetic Retinopathy test set.}
	\label{fig:DRexamples}
\end{figure}

\subsubsection{Adience}
Adience\footnote{http://www.openu.ac.il/home/hassner/Adience/data.html} dataset consists of $26580$ faces belonging to $2284$ subjects. We use the form of the dataset where faces have been pre-cropped and aligned. The dataset was preprocessed, using the methods described in a previous work~\cite{beckham2017unimodal}, so that the images are 256 pixels in width and height, and pixels values follow a $(0;1)$ normal distribution. The original dataset was split into five cross-validation folds. The training set consists of merging the first four folds which comprise a total of $15554$ images. From this, $10\%$ of the images are held out as part of a validation set. The last fold is used as test set. Some images of this dataset are shown in Figure \ref{fig:AdienceExamples}. Adience dataset has been used in other works for human age estimation but most of them solved the problem as a multi-class problem instead of using the ordinal relation between classes. Eidinger et al. \cite{eidinger2014age} presented an approach using support vector machines and neural networks. Chen et al. \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep CNNs. Levi and Hassner \cite{levi2015age} presented another convolutional network model for age estimation. As previously discussed, Beckham and Pal \cite{beckham2017unimodal} proposed a straightforward method to constrain discrete probability distributions to be unimodal.

\begin{figure}[!t]
	\centering
	\begin{tabular}{cccc}
		\includegraphics[width=2.4cm]{img/adience/1.jpg} & \includegraphics[width=2.4cm]{img/adience/2.jpg} & \includegraphics[width=2.4cm]{img/adience/3.jpg} & \includegraphics[width=2.4cm]{img/adience/4.jpg}\\
		\includegraphics[width=2.4cm]{img/adience/5.jpg} & \includegraphics[width=2.4cm]{img/adience/6.jpg} & \includegraphics[width=2.4cm]{img/adience/7.jpg} & \includegraphics[width=2.4cm]{img/adience/8.jpg}
	\end{tabular}						
	\caption{Examples taken from the Adience test set.}
	\label{fig:AdienceExamples}
\end{figure}

@@

\begin{table}[!t]
	\caption{Description of the architecture used for each dataset.}
	\label{table:Architecture}
	\centering
	\footnotesize
	\begin{tabular}{lclc}
		\hline\hline
		\multicolumn{2}{c}{\textbf{Diabetic Retinopathy}} & \multicolumn{2}{c}{\textbf{Adience}}  \\ \hline
		\textbf{Layer}  &         \textbf{Output}         & \textbf{Layer}      & \textbf{Output} \\ \hline
		2 Conv3x3@32s1  &           124x124x32            & Conv7x7@32s2        &   112x112x32    \\
		MaxPool2x2s2    &            62x62x32             & MaxPool3x3s2        &    55x55x32     \\
		2 Conv3x3@64s1  &            58x58x64             & 2 ResBlock3x3@64s1  &    55x55x32     \\
		MaxPool2x2s2    &            29x29x64             & ResBlock3x3@128s2   &    28x28x64     \\
		2 Conv3x3@128s1 &            25x25x128            & 2 ResBlock3x3@128s1 &    28x28x64     \\
		MaxPool2x2s2    &            12x12x128            & 1 ResBlock3x3@256s2 &    14x14x128    \\
		2 Conv3x3@128s1 &             8x8x128             & 2 ResBlock3x3@256s1 &    14x14x128    \\
		MaxPool2x2s2    &             4x4x128             & 1 ResBlock3x3@512s2 &     7x7x256     \\
		Conv4x4@128s1   &             1x1x128             & 2 ResBlock3x3@512s1 &     7x7x256     \\
		                &                                 & AveragePool7x7s2    &     1x1x256     \\ \hline\hline
	\end{tabular}	
\end{table}

@@

The performance gain of CLM over the nominal version reaches $16.8\%$ for DR and $11.9\%$ for Adience dataset. The improvement of the ordinal method for DR is higher than that for Adience. It seems that the method proposed in this work offers a more significant improvement as the given problem complexity increases. Moreover, when compared against alternative ordinal methods (many of them with a deep structure), CLMs are very competitive, possibly because of the flexibility provided by the threshold model structure (where the threshold of each class is independently adjusted).

\begin{table}[!t]
	\caption{Comparison between the best results of the CLM network, a nominal network (using softmax and cross-entropy) and previous works for both datasets.}
	\label{table:Comparison}
	\scriptsize
	\centering
	\def\arraystretch{1.3}
	\begin{tabular}{lccc}
		\hline\hline
		\multicolumn{4}{c}{\textbf{Diabetic Retinopathy}}\\
		\hline
		Method                                    & $\overline{QWK}_{(SD)}$ & $\overline{CCR}_{(SD)}$ & $\overline{\text{1-off}}_{(SD)}$ \\ \hline
		CLM network (proposal)                          &    $0.582_{(0.016)}$    &    $0.723_{(0.004)}$    &        $0.868_{(0.002)}$         \\
		Nominal network                           &    $0.498_{(0.013)}$    &    $0.692_{(0.014)}$    &        $0.854_{(0.006)}$         \\
		de la Torre et al. \cite{de2018weighted}     &  $0.537_{(\text{-})}$   &            -            &                -                 \\
		Nebot et al. \cite{nebot2016diabetic}  &  $0.555_{(\text{-})}$   &            -            &                -                 \\
		\hline
		\multicolumn{4}{c}{\textbf{Adience}}\\
		\hline
		Method                                    & $\overline{QWK}_{(SD)}$ & $\overline{CCR}_{(SD)}$ & $\overline{\text{1-off}}_{(SD)}$ \\ \hline
		CLM network (proposal)                          &    $0.881_{(0.005)}$    &    $0.519_{(0.013)}$    &        $0.894_{(0.005)}$         \\
		Nominal network                           &    $0.787_{(0.004)}$    &    $0.458_{(0.008)}$    &        $0.800_{(0.007)}$         \\
		Beckham and Pal \cite{beckham2017unimodal} &    $0.855_{(0.012)}$    &    $0.467_{(0.019)}$    &        $0.867_{(0.011)}$         \\
		Eidinger et al. \cite{eidinger2014age} &            -            &    $0.451_{(0.026)}$    &        $0.807_{(0.011)}$         \\
		Chen et al. \cite{chen2016cascaded} &            -            &    $0.529_{(0.060)}$    &        $0.885_{(0.022)}$         \\
		Levi and Hassner \cite{levi2015age}         &            -            &    $0.507_{(0.051)}$    &        $0.847_{(0.022)}$         \\ \hline\hline
	\end{tabular}
\end{table}

>>>>

\section{Introduction}
\label{sect:introduction}
Deep learning, introduced by Lecun~et al. \cite{lecun2015deep}, combines multiple machine learning techniques and allows computational models that are composed of numerous processing layers to learn representations of data with various levels of abstraction. These methods have dramatically improved the state-of-the-art in many domains, such as image classification~\cite{yu2017convolutional,fang2019image,zhu2019image}, speech recognition~\cite{song2018music,liu2018speech}, control problems~\cite{mnih2015human}, object detection~\cite{sun2018face,chu2018deep,olmos2018automatic}, privacy and security protection~\cite{tan2018deep,yuan2019adversarial}, recovery of human pose~\cite{hong2015multimodal}, semantic segmentation~\cite{li2019deep} and image retrieval~\cite{tzelepi2018deep,qayyum2017medical,bai2018optimization}. Convolutional Neural Networks (CNN) are one of the types of deep networks that are designed to process data that come in the form of multiple arrays. CNNs are appropriate for images, video, speech and audio processing, and they have been used extensively in the last years for automatic classification tasks~\cite{dong2014learning,ronneberger2015u}. For images, each colour channel is represented by a 2D array, and convolutional layers extract the main features from the pixels, and, after that, a fully connected layer classify every sample based on its extracted features. At the output of the CNN, a softmax function provides the probabilities of the set of classes predefined in the model for classification tasks. However, the softmax may not be the best option depending on the problem considered.

@@

\begin{figure}[!t]
	\centering
	\resizebox{\columnwidth}{!}{
	\begin{tabular}{cccc}
		\includegraphics[width=2.4cm]{img/retinopathy/1.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/2.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/3.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/4.jpg}\\
		\includegraphics[width=2.4cm]{img/retinopathy/5.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/6.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/7.jpg} & \includegraphics[width=2.4cm]{img/retinopathy/8.jpg}
	\end{tabular}}					
	\caption{Examples taken from the Diabetic Retinopathy test set.}
	\label{fig:DRexamples}
\end{figure}

\subsubsection{Adience}
Adience\footnote{http://www.openu.ac.il/home/hassner/Adience/data.html} dataset consists of $26580$ faces belonging to $2284$ subjects. We use the form of the dataset where faces have been pre-cropped and aligned. The dataset was preprocessed, using the methods described in a previous work~\cite{beckham2017unimodal}, so that the images are 256 pixels in width and height, and pixels values follow a $(0;1)$ normal distribution. The original dataset was split into five cross-validation folds. The training set consists of merging the first four folds which comprise a total of $15554$ images. From this, $10\%$ of the images are held out as part of a validation set. The last fold is used as test set. Some images of this dataset are shown in Figure \ref{fig:AdienceExamples}. Adience dataset has been used in other works for human age estimation but most of them solved the problem as a multi-class problem instead of using the ordinal relation between classes. Eidinger et al. \cite{eidinger2014age} presented an approach using support vector machines and neural networks. Chen et al. \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep CNNs. Levi and Hassner \cite{levi2015age} presented another convolutional network model for age estimation. As previously discussed, Beckham and Pal \cite{beckham2017unimodal} proposed a straightforward method to constrain discrete probability distributions to be unimodal. M. Duan et al. \cite{duan2018hybrid} proposed a hybrid approach that combines CNN with Extreme Learning Machine (ELM).

\begin{figure}[!t]
	\centering
	\resizebox{\columnwidth}{!}{
	\begin{tabular}{cccc}
		\includegraphics[width=2.4cm]{img/adience/1.jpg} & \includegraphics[width=2.4cm]{img/adience/2.jpg} & \includegraphics[width=2.4cm]{img/adience/3.jpg} & \includegraphics[width=2.4cm]{img/adience/4.jpg}\\
		\includegraphics[width=2.4cm]{img/adience/5.jpg} & \includegraphics[width=2.4cm]{img/adience/6.jpg} & \includegraphics[width=2.4cm]{img/adience/7.jpg} & \includegraphics[width=2.4cm]{img/adience/8.jpg}
	\end{tabular}}				
	\caption{Examples taken from the Adience test set.}
	\label{fig:AdienceExamples}
\end{figure}

@@

\begin{table}[!t]
	\caption{Description of the architecture used for each dataset.}
	\label{table:Architecture}
	\centering
	\footnotesize
	\resizebox{\columnwidth}{!}{
	\begin{tabular}{lclc}
		\hline\hline
		\multicolumn{2}{c}{\textbf{Diabetic Retinopathy}} & \multicolumn{2}{c}{\textbf{Adience}}  \\ \hline
		\textbf{Layer}  &         \textbf{Output}         & \textbf{Layer}      & \textbf{Output} \\ \hline
		2 Conv3x3@32s1  &           124x124x32            & Conv7x7@32s2        &   112x112x32    \\
		MaxPool2x2s2    &            62x62x32             & MaxPool3x3s2        &    55x55x32     \\
		2 Conv3x3@64s1  &            58x58x64             & 2 ResBlock3x3@64s1  &    55x55x32     \\
		MaxPool2x2s2    &            29x29x64             & ResBlock3x3@128s2   &    28x28x64     \\
		2 Conv3x3@128s1 &            25x25x128            & 2 ResBlock3x3@128s1 &    28x28x64     \\
		MaxPool2x2s2    &            12x12x128            & 1 ResBlock3x3@256s2 &    14x14x128    \\
		2 Conv3x3@128s1 &             8x8x128             & 2 ResBlock3x3@256s1 &    14x14x128    \\
		MaxPool2x2s2    &             4x4x128             & 1 ResBlock3x3@512s2 &     7x7x256     \\
		Conv4x4@128s1   &             1x1x128             & 2 ResBlock3x3@512s1 &     7x7x256     \\
		                &                                 & AveragePool7x7s2    &     1x1x256     \\ \hline\hline
	\end{tabular}}
\end{table}

@@

The performance gain of CLM over the nominal version reaches $16.8\%$ for DR and $11.9\%$ for Adience dataset. The DR dataset obtains a higher performance gain from the ordinal method than the Adience dataset. It seems that the method proposed in this work offers a more significant improvement as the given problem complexity increases. Moreover, when compared against alternative ordinal methods (many of them with a deep structure), CLMs are very competitive, possibly because of the flexibility provided by the threshold model structure (where the threshold of each class is independently adjusted). Our proposed model shows a high score and a very low standard deviation, which proves very high stability compared to other alternatives.

\begin{table}[!t]
	\caption{Comparison between the best results of the CLM network, a nominal network (using softmax and cross-entropy) and previous works for both datasets.}
	\label{table:Comparison}
	\scriptsize
	\centering
	\def\arraystretch{1.3}
	\begin{tabular}{lccc}
		\hline\hline
		\multicolumn{4}{c}{\textbf{Diabetic Retinopathy}}\\
		\hline
		Method                                    & $\overline{QWK}_{(SD)}$ & $\overline{CCR}_{(SD)}$ & $\overline{\text{1-off}}_{(SD)}$ \\ \hline
		CLM network (proposal)                          &    $0.582_{(0.016)}$    &    $0.723_{(0.004)}$    &        $0.868_{(0.002)}$         \\
		Nominal network                           &    $0.498_{(0.013)}$    &    $0.692_{(0.014)}$    &        $0.854_{(0.006)}$         \\
		de la Torre et al. \cite{de2018weighted}     &  $0.537_{(\text{-})}$   &            -            &                -                 \\
		Nebot et al. \cite{nebot2016diabetic}  &  $0.555_{(\text{-})}$   &            -            &                -                 \\
		\hline
		\multicolumn{4}{c}{\textbf{Adience}}\\
		\hline
		Method                                    & $\overline{QWK}_{(SD)}$ & $\overline{CCR}_{(SD)}$ & $\overline{\text{1-off}}_{(SD)}$ \\ \hline
		CLM network (proposal)                          &    $0.881_{(0.005)}$    &    $0.519_{(0.013)}$    &        $0.894_{(0.005)}$         \\
		Nominal network                           &    $0.787_{(0.004)}$    &    $0.458_{(0.008)}$    &        $0.800_{(0.007)}$         \\
		Beckham and Pal \cite{beckham2017unimodal} &    $0.855_{(0.012)}$    &    $0.467_{(0.019)}$    &        $0.867_{(0.011)}$         \\
		Eidinger et al. \cite{eidinger2014age} &            -            &    $0.451_{(0.026)}$    &        $0.807_{(0.011)}$         \\
		Chen et al. \cite{chen2016cascaded} &            -            &    $0.529_{(0.060)}$    &        $0.885_{(0.022)}$         \\
		Levi and Hassner \cite{levi2015age}         &            -            &    $0.507_{(0.051)}$    &        $0.847_{(0.022)}$         \\ 
		M. Duan et al. \cite{duan2018hybrid}         &            -            &    $0.523_{(0.057)}$    &        -         \\ \hline\hline
	\end{tabular}
\end{table}
